### 2020-06-01 report/sections/evalconcl/eval.tex
Report: update captionof's
Report: add protocol listings for forward-ref

\section{Binary Sessions: \fancyname{Calculator}}
\begin{itemize}
\item sample client implementation + snippet of generated code (full in appendix)
\item show that not all the UI is being controlled by the runtime (i.e. calculator app has a header/fake navbar)
\item show the calculator logic achievable using React Context
\item sample server implementation + snippet of generated code (full in appendix)
\end{itemize}

\section{Multiparty Sessions: \fancyname{Noughts and Crosses}}
\begin{itemize}
\item sample client implementation -- show the flexibility of factory API for sending to be used in the UI part (since move selection can be be made )
\end{itemize}

@@

>>>>

\section{Binary Sessions: \fancyname{Calculator}}
\label{section:evalcalculator}
\begin{itemize}
\item sample client implementation + snippet of generated code (full in appendix)
\item show that not all the UI is being controlled by the runtime (i.e. calculator app has a header/fake navbar)
\item show the calculator logic achievable using React Context
\item sample server implementation + snippet of generated code (full in appendix)
\end{itemize}

\begin{figure}[!ht]
\begin{lstlisting}[language=Scribble, tabsize=2]
module BinarySession;

global protocol Calculator(role Svr, role Client) {
  choice at Client {
    Add(number, number) from Client to Svr;
    Res(number)         from Svr to Client;
    do Calculator(Svr, Client);
  } or {
    Mult(number, number) from Client to Svr;
    Res(number)          from Svr to Client;
    do Calculator(Svr, Client);
  } or {
    Quit()               from Client to Svr;
    Terminate()          from Svr to Client;
  }
}
\end{lstlisting}
\captionof{lstlisting}{The \tprotocol{Calculator} Protocol}
\label{lst:calculator}
\end{figure}

\section{Multiparty Sessions: \fancyname{Noughts and Crosses}}
\label{section:evalgame}
\begin{itemize}
\item sample client implementation -- show the flexibility of factory API for sending to be used in the UI part (since move selection can be be made )
\end{itemize}

\begin{figure}[!ht]
\begin{lstlisting}[language=Scribble, tabsize=2]
module NoughtsAndCrosses;
type <typescript> "Coordinate" from "./Types" as Point;

global protocol Game(role Svr, role P1, role P2) {
	Pos(Point) from P1 to Svr;
	choice at Svr {
		Lose(Point)   from Svr to P2;
		Win(Point)    from Svr to P1;
	} or {
		Draw(Point)   from Svr to P2;
		Draw(Point)   from Svr to P1;
	} or {
		Update(Point) from Svr to P2;
		Update(Point) from Svr to P1;
		do Game(Svr, P2, P1);
	}
}
\end{lstlisting}
\captionof{lstlisting}{The \tprotocol{Noughts and Crosses} Protocol}
\label{lst:game}
\end{figure}

@@

\begin{figure}[!ht]
\begin{lstlisting}[language=Scribble, tabsize=2]
module ClientClient;

global protocol TwoBuyer(role A, role B, role S) {
	title(string)      from A to S;
	quote(number)      from S to A;
	quote(number)      from S to B;
	quoteByTwo(number) from A to B;
	choice at B {
		ok(string)     from B to S;
		date(string)   from S to B;
	} or {
		quit()         from B to S;
	}
}
\end{lstlisting}
\captionof{lstlisting}{The \tprotocol{Two Buyer} Protocol}
\label{lst:twobuyer}
\end{figure}

### 2020-06-01 report/sections/evalconcl/eval/benchmarks.tex
Report: update captionof's
Report: add protocol listings for forward-ref

\begin{figure}[!ht]
\begin{lstlisting}[language=Scribble]
global protocol PingPong(role Client, role Svr) {
	PING(number) from Client to Svr;
	choice at Svr {
		PONG(number) from Svr to Client;
		do PingPong(Client, Svr);
	} or {
		BYE(number) from Svr to Client;	
	}
}
\end{lstlisting}
\captionof{lstlisting}{Ping Pong Protocol}
\label{lst:pingpong}
\end{figure}

@@

We parameterise the Ping Pong protocol by $n > 0$, the number of
round-trip messages. This is standardised in the application logic
across experiments.
Upon establishing a connection, the experiment proceeds as follow:

@@

We note that the Ping Pong protocol implements a \textit{binary} session. 
It would be interesting to observe the overhead in a \textit{multiparty}
context, but due to limited time constraints,
our benchmarking suite does not support multiple browser targets.
Benchmarking multiparty protocols would also require writing multiple
distinct React applications using the generated APIs -- as this is currently
a manual process, doing this for multiple roles requires more time than
available.

@@

\begin{figure}[!ht]
\centering
\includegraphics[width=0.5\textwidth]{pingpongclient}
\captionof{figure}{User interface of Ping Pong \trole{Client}}
\label{fig:pingpongclient}
\end{figure}

\subparagraph{Ping Pong \trole{Client} on React:}
\begin{itemize}

@@

\subparagraph{Ping Pong \trole{Svr} on Node:}
\begin{itemize}

@@

The benchmark compares three implementations of the Ping Pong protocol:

\subparagraph{bare:}
The \texttt{bare} implementation directly interfaces with 
WebSocket primitives for sending and receiving. 
The implementation executes the Ping Pong protocol, but does not 
guarantee communication safety by construction -- e.g. the user can click
the \tmsg{PING} button multiple times before a \tmsg{PONG} message 
is received, violating channel linearity. 
This represents the typical developer implementation without using the
MPST framework.

@@

render() {
	return this.state.visible 
		? <button
		   ref={this.state.button}
		   onClick={this.click.bind(this)}
		   >Ping</button>
		: <button>Ping</button>;
}
\end{lstlisting}
\captionof{lstlisting}{Preventing user-triggered linearity violation in
\texttt{bare_safe} Ping Pong}
\label{lst:workaround}
\end{figure}

The workaround is tailored to the Ping Pong protocol
and toggles the \texttt{visible} flag to hide the button that triggers
a send action when \trole{Client} transitions to a receive state.
In fact, this ``workaround'' precisely generalises to our 
\texttt{mpst} implementation of having some form of wrapper
component that renders the UI component corresponding to the
current EFSM state. Our API generation strategy formally implements
the EFSM for browser-side logic.

@@

\begin{figure}[!ht]
\centering
\begin{subfigure}[b]{\textwidth}
\centering
\begin{lstlisting}[language=javascript,tabsize=2]
socket.onmessage = ({ data }) => {
	const { label, payload } = JSON.parse(data.toString());
	if (label === 'PING') {
		let count: number = payload[0];
		console.timeLog(LABEL, ++count);
		if (count === MSGS) {
			socket.send(JSON.stringify({
				label: 'BYE',
				payload: [count],
			}));
		} else {
			socket.send(JSON.stringify({
				label: 'PONG',
				payload: [count],
			}));
		}
	} else {
		throw new Error(`Unrecognised label: ${label}`);
	}
}
\end{lstlisting}
\caption{\texttt{bare} implementation of Ping Pong \trole{Svr}}
\label{lst:nodepingpongbare}
\end{subfigure}
\hfill
\begin{subfigure}[b]{\textwidth}
\centering
\begin{lstlisting}[language=javascript,tabsize=2]
// NOTE: `Implementation' namespace omitted for brevity.

const logic = new S14({
	PING: (count) => {
		console.timeLog(LABEL, ++count);
		if (count === MSGS) {
			return new S16([Labels.S16.BYE, [count], new S15()]);
		} else {
			return new S16([Labels.S16.PONG, [count], logic]);
		}
 	}
});
\end{lstlisting}
\caption{\texttt{mpst} implementation of Ping Pong \trole{Svr}}
\label{lst:nodepingpongmpst}
\end{subfigure}
\captionof{figure}{Comparison of \texttt{bare} and \texttt{mpst}
implementations for Ping Pong \trole{Svr}}
\label{fig:nodepingpong}
\end{figure}

>>>>

\begin{figure}[!ht]
\begin{lstlisting}[language=Scribble]
global protocol PingPong(role Client, role Svr) {
	PING(number) from Client to Svr;
	choice at Svr {
		PONG(number) from Svr to Client;
		do PingPong(Client, Svr);
	} or {
		BYE(number) from Svr to Client;	
	}
}
\end{lstlisting}
\captionof{lstlisting}{The \tprotocol{Ping Pong} Protocol}
\label{lst:pingpong}
\end{figure}

@@

We parameterise the \tprotocol{Ping Pong} protocol by $n > 0$, the number of
round-trip messages. This is standardised in the application logic
across experiments.
Upon establishing a connection, the experiment proceeds as follow:

@@

We note that the \tprotocol{Ping Pong} protocol 
implements a \textit{binary} session. 
It would be interesting to observe the overhead in a \textit{multiparty}
context, but due to limited time constraints,
our benchmarking suite does not support multiple browser targets.
Benchmarking multiparty protocols would also require writing multiple
distinct React applications using the generated APIs -- as this is currently
a manual process, doing this for multiple roles requires more time than
available.

@@

\begin{figure}[!ht]
\centering
\includegraphics[width=0.5\textwidth]{pingpongclient}
\captionof{figure}{User interface of \tprotocol{Ping Pong} \trole{Client}}
\label{fig:pingpongclient}
\end{figure}

\subparagraph{\tprotocol{Ping Pong} \trole{Client} on React:}
\begin{itemize}

@@

\subparagraph{\tprotocol{Ping Pong} \trole{Svr} on Node:}
\begin{itemize}

@@

The benchmark compares three implementations of the \tprotocol{Ping Pong} protocol:

\subparagraph{bare:}
The \texttt{bare} implementation directly interfaces with 
WebSocket primitives for sending and receiving. 
The implementation executes the \tprotocol{Ping Pong} protocol, but does not 
guarantee communication safety by construction -- e.g. the user can click
the \tmsg{PING} button multiple times before a \tmsg{PONG} message 
is received, violating channel linearity. 
This represents the typical developer implementation without using the
MPST framework.

@@

render() {
	return this.state.visible 
		? <button
		   ref={this.state.button}
		   onClick={this.click.bind(this)}
		   >Ping</button>
		: <button>Ping</button>;
}
\end{lstlisting}
\captionof{lstlisting}{Preventing user-triggered linearity violation in
\texttt{bare_safe} \tprotocol{Ping Pong}}
\label{lst:workaround}
\end{figure}

The workaround is tailored to the \tprotocol{Ping Pong} protocol
and toggles the \texttt{visible} flag to hide the button that triggers
a send action when \trole{Client} transitions to a receive state.
In fact, this ``workaround'' precisely generalises to our 
\texttt{mpst} implementation of having some form of wrapper
component that renders the UI component corresponding to the
current EFSM state. Our API generation strategy formally implements
the EFSM for browser-side logic.

@@

\begin{figure}[!ht]
\centering
\begin{subfigure}[b]{\textwidth}
\centering
\begin{lstlisting}[language=javascript,tabsize=2]
socket.onmessage = ({ data }) => {
	const { label, payload } = JSON.parse(data.toString());
	if (label === 'PING') {
		let count: number = payload[0];
		console.timeLog(LABEL, ++count);
		if (count === MSGS) {
			socket.send(JSON.stringify({
				label: 'BYE',
				payload: [count],
			}));
		} else {
			socket.send(JSON.stringify({
				label: 'PONG',
				payload: [count],
			}));
		}
	} else {
		throw new Error(`Unrecognised label: ${label}`);
	}
}
\end{lstlisting}
\caption{\texttt{bare} implementation of \tprotocol{Ping Pong} \trole{Svr}}
\label{lst:nodepingpongbare}
\end{subfigure}
\hfill
\begin{subfigure}[b]{\textwidth}
\centering
\begin{lstlisting}[language=javascript,tabsize=2]
// NOTE: `Implementation' namespace omitted for brevity.

const logic = new S14({
	PING: (count) => {
		console.timeLog(LABEL, ++count);
		if (count === MSGS) {
			return new S16([Labels.S16.BYE, [count], new S15()]);
		} else {
			return new S16([Labels.S16.PONG, [count], logic]);
		}
 	}
});
\end{lstlisting}
\caption{\texttt{mpst} implementation of \tprotocol{Ping Pong} \trole{Svr}}
\label{lst:nodepingpongmpst}
\end{subfigure}
\captionof{figure}{Comparison of \texttt{bare} and \texttt{mpst}
implementations for \tprotocol{Ping Pong} \trole{Svr}}
\label{fig:nodepingpong}
\end{figure}
